Smart Model Elimination for Efficient Automated Machine Learning

### 2024-06-16 paper.tex
another dataset

\subsection{Overview}
When considering a specific dataset, various aspects regarding the distribution and relationship within the data can give us clues as to which models are likely to perform best. For example, linear models tend to work better on datasets where there is a linear relationship between features. As a result, we can develop a boosting model to predict the best models based off extracted attributes from a dataset.

\begin{table}
  \caption{Model Options}
  \label{model-options-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    Attribute & Model & Purpose                         \\
    \midrule
    \# of Rows & $|\mathbf{A}|_r$ & Data Size \\
    \hline
    \# of Columns & $|\mathbf{A}|_c$ & Data Size \\
    \hline
    Target Distribution &
    $\frac{\max(\text{count}_1, \text{count}_0)}{|\mathbf{A}|_r}$
    & Data Type \\
    \hline
    Numerical Columns &
    $\frac{|\mathbf{A}_{num}|}{|\mathbf{A}|_{c}}$
    & Data Type \\
    \hline
    Binary Categorical Columns &
    $\frac{|\mathbf{A}_{bin}|}{|\mathbf{A}|_{c}}$
    & Data Type \\
    \hline
    Mean \# of Distinct Values  & $\mu \left\{\text{\# of distinct values in column}\right\}$ & Data Type \\ 
    \hline
    Mean IQR  & $\mu \left\{\text{IQR of column}\right\}$ & Data Distribution \\ 
    \hline
    Standard Deviation of IQR  & $\sigma \left\{\text{IQR of column}\right\}$ & Data Distribution \\
    \hline
    Mean Q1 & $\mu \left\{\text{Q1 of column}\right\}$ & Data Distribution \\
    \hline
    Standard Deviation of Q1  & $\sigma \left\{\text{Q1 of column}\right\}$ & Data Distribution \\
    \hline
    Mean Q3  & $\mu \left\{\text{Q3 of column}\right\}$ & Data Distribution \\
    \hline
    Standard Deviation of Q3  & $\sigma \left\{\text{Q3 of column}\right\}$ & Data Distribution \\
    \hline
    Mean Percent of Z-Score Outliers  & 
    $\mu \left\{ \frac{\text{\# of Z-Score Outliers}}{|\mathbf{A}|_r} \right\}$ & Outlier Data Points \\
    \hline
    Standard Deviation of Percent of Z-Score Outliers  &
    $\sigma \left\{ \frac{\text{\# of Z-Score Outliers}}{|\mathbf{A}|_r} \right\}$ & Outlier Data Points \\
    \hline
    Mean Correlation& $\mu \left\{\text{Correlation Matrix}\right\}$ & Data linearity \\
    \hline
    Standard Deviation of Correlation & $\sigma \left\{\text{Correlation Matrix}\right\}$ & Data Linearity \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

\subsection{Observation}
When considering a specific dataset, various aspects regarding the distribution and relationship within the data can give us clues as to which models are likely to perform best. For example, linear models tend to work better on datasets where there is a linear relationship between features. As a result, we can develop a boosting model to predict the best models based off extracted attributes from a dataset.

\begin{table}
  \caption{Attributes extracted from a dataset}
  \label{model-options-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    Attribute & Formula & Purpose\\
    \midrule
    \# of Rows & $|\mathbf{A}|_r$ & Data Size \\
    \hline
    \# of Columns & $|\mathbf{A}|_c$ & Data Size \\
    \hline
    Target Distribution &
    $\frac{\max(\text{count}_1, \text{count}_0)}{|\mathbf{A}|_r}$
    & Data Type \\
    \hline
    Numerical Columns &
    $\frac{|\mathbf{A}_{num}|}{|\mathbf{A}|_{c}}$
    & Data Type \\
    \hline
    Binary Categorical Columns &
    $\frac{|\mathbf{A}_{bin}|}{|\mathbf{A}|_{c}}$
    & Data Type \\
    \hline
    Mean \# of Distinct Values  & $\mu \left\{\text{\# of distinct values in column}\right\}$ & Data Type \\ 
    \hline
    Mean IQR  & $\mu \left\{\text{IQR of column}\right\}$ & Data Distribution \\ 
    \hline
    Standard Deviation of IQR  & $\sigma \left\{\text{IQR of column}\right\}$ & Data Distribution \\
    \hline
    Mean Q1 & $\mu \left\{\text{Q1 of column}\right\}$ & Data Distribution \\
    \hline
    Standard Deviation of Q1  & $\sigma \left\{\text{Q1 of column}\right\}$ & Data Distribution \\
    \hline
    Mean Q3  & $\mu \left\{\text{Q3 of column}\right\}$ & Data Distribution \\
    \hline
    Standard Deviation of Q3  & $\sigma \left\{\text{Q3 of column}\right\}$ & Data Distribution \\
    \hline
    Mean Percent of Z-Score Outliers  & 
    $\mu \left\{ \frac{\text{\# of Z-Score Outliers}}{|\mathbf{A}|_r} \right\}$ & Outlier Data Points \\
    \hline
    Standard Deviation of Percent of Z-Score Outliers  &
    $\sigma \left\{ \frac{\text{\# of Z-Score Outliers}}{|\mathbf{A}|_r} \right\}$ & Outlier Data Points \\
    \hline
    Mean Correlation& $\mu \left\{\text{Correlation Matrix}\right\}$ & Data linearity \\
    \hline
    Standard Deviation of Correlation & $\sigma \left\{\text{Correlation Matrix}\right\}$ & Data Linearity \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Data Extraction}
To train our boosting model, we collected over a thousand datasets from kaggle, a popular website for hosting public datasets. After used various medical terms to filter down datasets and limiting our search to tabular, binary-classification datasets, we were left with 268 datasets. We extracted the attributes in Table 2. from each dataset and ran every model on each dataset to get the performace of each model. We trained a multilabel regression model on these attributes to predict the rank of every model on a specific dataset.
